Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2019-11-07 manuscript/climate_sensitivity_draft.Rmd
added mean R values for drought, adjusted discussion of drought strength
finishes issue #60

Community-level tree growth responses to these droughts were modest, with modal resistance values of **#, #, and #** for 1966, 1977, and 1999, respectively (Fig. 1b). In each drought, roughly 30% of the cored trees suffered $\ge$ 30% growth reductions ($R$ $\le$ 0.7): 29.43% in 1966, 31.86% in 1977, and 26.81% in 1999. Some trees exhibited increased growth: ($R$ > 1.0): 26.31% in 1966, 21.92% in 1977, and 25.57% in 1999.

@@

Although the physiological mechanisms discussed above lead to generally consistent directions of growth responses to tree height and hydaulic traits across droughts, indicating the universality of the underlying mechanisms, the relative importantance of the drivers varied widely across droughts, indicating an interaction with drought characteristics (Tables 4-5). Although there were not signficant differences in $R$ among the drought years, $R$ tended to be somewhat lower in 1966 (Fig. 1, Table 4). Height and dominant canopy position had the stronger negative effects in this drought than in the others (Tables 4-5), consistent with the finding that height becomes a stronger predictor of mortality as the severity of the drought increases [@stovall_tree_2019]. In 1977, which was the weakest drought (Fig. S3, Table S2), ring-porous species had lower $R$, contrasting with their higher $R$ in the other two droughts and perhaps indicating that ring-porosity provides less of an advantage in terms of drought adaptation under less extreme droughts, consistent with @elliott_forest_2015. Somewhat surprisingly, but consistent with the weakening drought sensitivity observed by [@helcoski_growing_2019], the highest overall $R$ was observed during the 1999 drought, which was the most intense and second-longest in duration. For this drought, xylem architecture was the most useful predictor, with diffuse-porous species suffering most. Thus, responses differed with the nature of each drought.  Of course, site characteristics also define the nature of droughts, and comparisons of size and trait effects across sites--and across more droughts at a single site--would be of great value to elucidating the mechanisms through which drought characteristics interact with driver variables to shape tree growth responses.

>>>>

Community-level tree growth responses to these droughts were modest, with mean resistance values of 0.86, 0.84, and 0.86 for 1966, 1977, and 1999, respectively (Fig. 1b). In each drought, roughly 30% of the cored trees suffered $\ge$ 30% growth reductions ($R$ $\le$ 0.7): 29.43% in 1966, 31.86% in 1977, and 26.81% in 1999. Some trees exhibited increased growth: ($R$ > 1.0): 26.31% in 1966, 21.92% in 1977, and 25.57% in 1999.

@@

Although the physiological mechanisms discussed above lead to generally consistent directions of growth responses to tree height and hydaulic traits across droughts, indicating the universality of the underlying mechanisms, the relative importantance of the drivers varied widely across droughts, indicating an interaction with drought characteristics (Tables 4-5). Although there were not signficant differences in $R$ among the drought years, $R$ tended to be somewhat lower in 1966 (Fig. 1, Table 4). Height and dominant canopy position had the stronger negative effects in this drought than in the others (Tables 4-5), which could potentially be linked to the fact that this was the longest-lasting drought and and that height becomes a stronger predictor of mortality as the severity of the drought increases [@stovall_tree_2019]. In 1977, which was the weakest drought in terms of PDSI (Fig. S2, Table S3), ring-porous species had lower $R$, contrasting with their higher $R$ in the other two droughts and perhaps indicating that ring-porosity provides less of an advantage in terms of drought adaptation under less extreme droughts, consistent with @elliott_forest_2015. In the 1999 drought, which was the most intense in terms of growing season PDSI and the second-longest in duration, xylem architecture was the most useful predictor, with diffuse-porous species suffering most. Thus, responses differed with the nature of each drought. Of course, site characteristics also define the nature of droughts, and comparisons of size and trait effects across sites--and across more droughts at a single site--would be of great value to elucidating the mechanisms through which drought characteristics interact with driver variables to shape tree growth responses.
